### 2024-01-17 outputs/paper/paper.qmd
added starter files

---
title: "Hannah Yu"
subtitle: "My subtitle if needed"
author: 
  - First author
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

You can and should cross-reference sections and sub-sections.

The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("outputs/models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References

### 2024-01-18 scripts/01-download_data.R
Downloaded the raw_homicide_data from Open Data Toronto

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Author: Hannah Yu [...UPDATE THIS...]
+# Date: 17 January 2024 [...UPDATE THIS...]
+# Contact: realhannah.yu@mail.utoronto.ca [...UPDATE THIS...]
 # License: MIT
 # Pre-requisites: [...UPDATE THIS...]
 # Any other information needed? [...UPDATE THIS...]
@@ -11,16 +11,29 @@
 #### Workspace setup ####
 library(opendatatoronto)
 library(tidyverse)
+library(dplyr)
 # [...UPDATE THIS...]
 
 #### Download data ####
 # [...ADD CODE HERE TO DOWNLOAD...]
 
+raw_homicide_data <-
+
+  list_package_resources("7d72bbbe-8adc-4b36-8ad1-5359f1c7a9cc") |>
+  filter(format == "CSV") |>
+  head(1) |>
+  get_resource()
 
 
 #### Save data ####
 # [...UPDATE THIS...]
 # change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
+#write_csv(the_raw_data, "inputs/data/raw_data.csv") 
 
-         
+write_csv(
+  x = raw_homicide_data,
+  file = "inputs/data/raw_homicide_data.csv"
+)
+
+        
+

### 2024-01-18 scripts/00-simulate_data.R
Simulated Data for Homicide Data

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Author: Hannah Yu [...UPDATE THIS...]
+# Date: 17 January 2024 [...UPDATE THIS...]
+# Contact: realhannah.yu@mail.utoronto.ca [...UPDATE THIS...]
 # License: MIT
 # Pre-requisites: [...UPDATE THIS...]
 # Any other information needed? [...UPDATE THIS...]
@@ -10,10 +10,75 @@
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+library(knitr)
+library(janitor)
+library(lubridate)
+library(opendatatoronto)
 
 #### Simulate data ####
-# [...ADD CODE HERE...]
+
+#### Data expectations ####
+# There are a similar number of subdivisions per ward
+# Voter turnout varies by ward
+# Wards with more subdivisions will have a higher turnout
+
+#### Simulate voter turnout by ward ####
+#based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html#simulate
+
+set.seed(853)
+
+num_obs <- 1398
+
+simulated_data <- 
+  tibble(
+    year = sample(2004:2023, num_obs, replace = TRUE),
+    month = sample(1:12, num_obs, replace = TRUE),
+    day = sample(1:31, num_obs, replace = TRUE),
+    weekday = sample(c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday",
+                       "Friday", "Saturday"), num_obs, replace = TRUE),
+    homicide_type = sample(c("Shooting", "Stabbing", "Other"),  num_obs,
+                           replace = TRUE),
+    neighborhood_number = sample(1:158, num_obs, replace = TRUE),
+)
 
 
 
+# Print the first few rows of the tibble
+print(simulated_data)
+
+
+#Bar Plot for Homicide Type
+library(ggplot2)
+
+
+ggplot(simulated_data, aes(x = homicide_type, fill = homicide_type)) +
+  geom_bar() +
+  labs(title = "Distribution of Homicide Types", x = "Homicide Type",
+       y = "Count") +
+  theme_minimal()
+
+#Histogram for Neighborhood Number
+
[…]

### 2024-01-18 scripts/01-download_data.R

@@ -35,5 +35,5 @@ write_csv(
   file = "inputs/data/raw_homicide_data.csv"
 )
 
-        
+
 

### 2024-01-18 scripts/00-simulate_data.R
Added descriptions

@@ -18,11 +18,13 @@ library(opendatatoronto)
 #### Simulate data ####
 
 #### Data expectations ####
-# There are a similar number of subdivisions per ward
-# Voter turnout varies by ward
-# Wards with more subdivisions will have a higher turnout
+# The year, month, and day variables contain values within their specified ranges.
+# The weekday variable only include the days of the week from Monday to Sunday.
+# The homicide_type variable only include the following categories: "Shooting," "Stabbing," and "Other." 
+# The neighborhood_number variable include integers ranging from 1 to 158.
+# Columns: year, month, day, weekday, homicide_type, neighborhood_number
 
-#### Simulate voter turnout by ward ####
+#### Simulate Toronto Homicide Data ####
 #based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html#simulate
 
 set.seed(853)

### 2024-01-19 scripts/02-data_cleaning.R
trial

@@ -1,8 +1,8 @@
 #### Preamble ####
 # Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Author: Hannah Yu
+# Date: 19 January 2023  
+# Contact: realhannah.yu@mail.utoronto.ca 
 # License: MIT
 # Pre-requisites: [...UPDATE THIS...]
 # Any other information needed? [...UPDATE THIS...]

### 2024-01-19 scripts/02-data_cleaning.R
trial

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
+# Purpose: Cleans the raw plane data recorded by two observers.....  
 # Author: Hannah Yu
 # Date: 19 January 2023  
 # Contact: realhannah.yu@mail.utoronto.ca 
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites:  
+# Any other information needed? 
 
 #### Workspace setup ####
 library(tidyverse)

### 2024-01-19 scripts/02-data_cleaning.R
cleaned dataset and created new files

@@ -1,44 +1,103 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers.....  
+# Purpose: Cleans the raw homicide data by and selecting relevant columns. 
 # Author: Hannah Yu
 # Date: 19 January 2023  
 # Contact: realhannah.yu@mail.utoronto.ca 
 # License: MIT
-# Pre-requisites:  
-# Any other information needed? 
+# Pre-requisites: 01-download_data.R
+
 
 #### Workspace setup ####
 library(tidyverse)
+library(janitor)
+library(dplyr)
+library(readr)
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+raw_data <- read_csv("inputs/data/raw_homicide_data.csv")
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
+#### basic cleaning ####
+
+# based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html
+raw_homicide_data <-
+  read_csv(
+    file = "inputs/data/raw_homicide_data.csv",
+    show_col_type = FALSE
+  )
+
+# Make the names easier to type 
+cleaned_homicide_data <-
+  clean_names(raw_homicide_data)
+
+# Have a look at the first six rows
+head(cleaned_homicide_data)
+
+
+# select columns of interest #
+cleaned_homicide_data <-
[…]

### 2024-01-20 outputs/paper/paper.qmd
added new variable, started paper

@@
 author: 
-  - First author
-thanks: "Code and data are available at: LINK."
+  - Hannah Yu
+thanks: "Code and data are available at: ."
 date: today
@@
 
+#### Workspace set-up ####
+
+## Load packages
 library(tidyverse)
 library(palmerpenguins)
+library(ggplot2)
+
+#read in cleaned data #
+cleaned_homicide_data <- read.csv(here::here("inputs/data/cleaned_homicide_data.csv"))
+show_col_types = FALSE
 ```
@@
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+ggplot(cleaned_homicide_data, aes(x = homicide_type)) +
+  geom_bar(fill = "skyblue", color = "black") +
+  labs(title = "Distribution of Homicide Types",
+       x = "Homicide Types",
+       y = "Count") +
+  theme_minimal()
+
+# ggplot(penguins, aes(x = island, fill = species)) +
+#   geom_bar(alpha = 0.8) +
+#   scale_fill_manual(values = c("darkorange","purple","cyan4"),
+#                     guide = "none") +
+#   theme_minimal() +
+#   facet_wrap(~species, ncol = 1) +
+#   coord_flip()
+```
+
+```{r}
+ggplot(cleaned_homicide_data, aes(x = occurrence_year, y = count, group = 1)) +
+  geom_line(color = "blue") +
+  geom_point(color = "blue") +
+  labs(title = "Homicide Occurrences Over the Years",
+       x = "Year",
+       y = "Number of Homicides") +
+  theme_minimal()
 ```

### 2024-01-20 scripts/02-data_cleaning.R

@@ -61,7 +61,13 @@ cleaned_homicide_data <-
            between(lubridate::month(occurrence_date), 9, 11) ~ "Fall",
            TRUE ~ "Winter")
   )
-    
+
+# Count the occurrence of homicides by year #
+cleaned_homicide_data <- 
+  cleaned_homicide_data |>
+  group_by(occurrence_year)|>
+  mutate(homicide_count = n())
+
 
 #### Save data ####
 write_csv(

### 2024-01-20 scripts/02-data_cleaning.R
homicide count

@@ -67,6 +67,7 @@ cleaned_homicide_data <-
   cleaned_homicide_data |>
   group_by(occurrence_year)|>
   mutate(homicide_count = n())
+head(cleaned_homicide_data)
 
 
 #### Save data ####

### 2024-01-23 outputs/paper/paper.qmd
lots of changes

@@
 library(tidyverse)
-library(palmerpenguins)
 library(ggplot2)
+library(kableExtra)
 
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+To investigate the homicide cases in Toronto, I obtained the dataset “ Police Annual Statistical Report - Homicides (@citehomicidedata) from the Toronto Open Data Portal (@ropendatatoronto).
+Data was cleaned and analyzed using the open source statistical programming language R [@citer], with additional support from 'tidyverse' [@rtidyverse], 'ggplot2' [@rggplot2], 'kableExtra' [@rkableExtra],  Further details about data extraction and analyzation processes will be discussed in the following sections below.
+
+This dataset includes a thorough list of alongside every homicide case in Toronto from the year 2004-2020.
+We cleaned the dataset and selected variables that are relevant to our analysis of homicide cases.
+Our variable of interest includes the year the offence occurred, police division where the offence occurred, type of homicide (shooting, stabbing, other), date of offence, and identifier of neighbourhood using Toronto’s 158 neighbourhood structure.
+In addition to these variables, we also constructed three more variables: day of the week when the offence occurred (Monday-Sunday), weekend?
+season when the offence occurred (fall, winter, spring, summer), and count of homicide cases by year to further aid our data analysis (See summary statistics).
+This dataset was last refreshed on January 11, 2024.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| message: false
 #| echo: false
+#| label: tbl-cleaned_data
+#| tbl-cap: Sample of Cleaned Homicide Data
+
+head (cleaned_homicide_data, 5) |>
+  kable(
+    col.names = c("Year", "Division", "Homicide Type", "Date", "Hood ID", "Day of the Week", "Season", "Homicide Count"),
+  booktabs = TRUE
+  )
+
+```
+We listed out the number of homicide cases each year in Table 2 and created the summary statistics table of the homicide count per year.
+We found the year 2011 has the least cases of homicide with only 51 cases while the year 2018 with the total of 97 cases has the most number of homicide cases.
+
+```{r}
+#| message: false
+#| echo: false
+#| label: tbl-homicide_count
+#| tbl-cap: Total Homicide Cases Each Year
+
+# Assuming 'occurrence_year' is the column containing the year information
+homicide_counts_by_year <- cleaned_homicide_data |>
+  group_by(occurrence_year) |>
+  summarise(total_cases = n())
+#installed.packages("git")
+library(gt)
+
+homicide_counts_by_year %>%
+  gt() %>%
+  tab_header(
+    title = "Total Homicide Cases Each Year"
+  ) %>%
+  fmt_number(
+    columns = vars(total_cases),
+    decimals = 0
+  )
+
+# # Assuming 'occurrence_year' is the column containing the year information
+# homicide_counts_by_year <- cleaned_homicide_data |>
+#   group_by(occurrence_year) |>
+#   summarise(total_cases = n())
+# 
+# # Print the resulting table
+# kable(homicide_counts_by_year, caption = "Total Homicide Cases Each Year")
+
+```
+Of all the homicide cases, the cases caused by shooting is significantly higher than other type of homicides.
+In the years of our examination, shooting takes up more than 600 cases of homicides while stabbing takes over more than 250, and all other types together takes up around 300.
+
+```{r,out.width='20%', echo=FALSE}
+#| message: false
+#| echo: false
+#| label: fig - distribution_by_type
+#| fig-cap: Distribution of Homicide Types
 
@@
 
-# ggplot(penguins, aes(x = island, fill = species)) +
-#   geom_bar(alpha = 0.8) +
-#   scale_fill_manual(values = c("darkorange","purple","cyan4"),
-#                     guide = "none") +
-#   theme_minimal() +
-#   facet_wrap(~species, ncol = 1) +
-#   coord_flip()
 ```
+We then shift our attention to the homicide trend over time.
+Figure 2 portrays the trend of homicide cases throughout the span of almost 2 decades.
+With the obvious fluctuations, We do not observe a clear increase or decrease of cases across time.
+However, the homicide cases are decreasing after it peaked in 2018.
 
+Following the grand picture across time, Figure 3 breaks down the homicide cases over time into further details based on the type of homicide.
+The trend of the three types of homicide cases generally follow the overall trend in figure 2.
+Shooting always prevail over other types of homicide cases even in its lowest year.
+We conclude that there is no clear increase or decrease of homicide across time.
 ```{r}
-ggplot(cleaned_homicide_data, aes(x = occurrence_year, y = count, group = 1)) +
+#| message: false
+#| echo: false
+#| label: fig-occurrence
+#| fig-cap: Homicide Occurrences Over Years
+summarized_homicide_data <- cleaned_homicide_data |>
+  group_by(occurrence_year) |>
+  summarise(homicide_count = n())
+
+ggplot(cleaned_homicide_data, aes(x = occurrence_year, y = homicide_count, group = 1)) +
   geom_line(color = "blue") +
   geom_point(color = "blue") +
-  labs(title = "Homicide Occurrences Over the Years",
+  labs(title = "Homicide Occurrences Over Years",
        x = "Year",
@@
 ```
+```{r}
+#| message: false
+#| echo: false
+#| label: fig-specific_occurrence
+#| fig-cap: Total Homicide Cases Over Years by Homicide Type
 
-Talk more about it.
+count_by_type <- cleaned_homicide_data |>
+  group_by(occurrence_year, homicide_type) |>
+  summarise(total_cases = n(), .groups = 'drop_last')
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+ggplot(count_by_type,
+       aes(x = occurrence_year, y = total_cases, color = homicide_type)) +
+  geom_line() + 
+  labs(title = "Total Homicide Cases Over Years by Homicide Type",
+       x = "Year",
+       y = "Total Cases",
+       color = "Homicide Type") +
+  theme_minimal()
+
+```
+
+We want to determine whether if there are a specific time in the year when homicides are more likely to occurr.
+To investigate our interest, we created figure 4 that displays the homicide counts by every day of the week.
+Sunday with a total of 200 cases is the day with the most cases following by Saturday and Friday with 197 and 175 respectively.
+Thursday is the day with the least cases with 124 cases.
+We conclude that homicides are more likely to occur in weekends.
 
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
+#| message: false
 #| echo: false
-#| warning: false
-#| message: false
+#| include: false
+#| label: tbl-casebyweek
+#| tbl-cap: Case_by_week
 
-analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))
+homicide_counts_by_day <- cleaned_homicide_data %>%
+  group_by(day_of_week) %>%
+  summarise(total_cases = n())
 
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
+print(homicide_counts_by_day)
+
 ```
 
-Talk way more about it.
+```{r}
+#| message: false
+#| echo: false
+#| label: fig-case_by_week1
+#| fig-cap: Distribution of Homicide Cases by Day of the Week
 
-# Model
+cleaned_homicide_data$day_of_week <- factor(cleaned_homicide_data$day_of_week, 
+                                             levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
 
-The goal of our modelling strategy is twofold. Firstly,...
+ggplot(cleaned_homicide_data, aes(x = day_of_week, fill = day_of_week)) +
+  geom_bar() +
+  labs(title = "Distribution of Homicide Cases by Day of the Week",
+       x = "Day of the Week",
+       y = "Number of Homicides") +
+  theme_minimal()
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+```
 
-## Model set-up
+We perform similar analysis with the count of homicide cases by season and found that the seasonal data are distributed somewhat evenly with fall having slightly more cases than other weather but the differnece is to small and not worth noting.
+```{r}
+#| message: false
+#| echo: false
+#| label: tbl-case_by_season
+#| tbl-cap: Distribution of Homicide Cases by Season
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+cleaned_homicide_data$season <- factor(cleaned_homicide_data$season, 
+                                             levels = c("Fall", "Winter", "Spring", "Summer"))
+ggplot(cleaned_homicide_data, aes(x = factor(season), fill = season)) +
+  geom_bar() +
+  labs(title = "Distribution of Homicide Cases by Season",
+       x = "Season",
+       y = "Number of Homicides") +
+  theme_minimal()
+```
+To investigate the relationship between geographical location and homicide cases, we constructed figure .
+This bar plot illustrates the distribution of homicide cases in Toronto's 140 neighborhoods.
+This visualization enables the identification of patterns or areas of concerns in the city.
+While most neighborhoods have fewer than 10 cases, we notice that several neighborhoods stand out with significantly higher homicide incidences compared to the rest.
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+```{r}
+#| message: false
+#| echo: false
+#| label: fig-neighborhoods_distribution
+#| fig-cap: Distribution of Homicide in Neighborhoods
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+nei<- cleaned_homicide_data %>%
+  group_by(hood_id) %>%
+  summarise(total_cases = n())
 
-### Model justification
+ggplot(cleaned_homicide_data, aes(x = hood_id)) +
+  geom_bar(fill = "skyblue", color = "black") +
+  labs(title = "Distribution of Homicide Cases in 140 Neighborhoods",
+       x = "Neighborhood Number",
+       y = "Number of Homicide Cases") +
+  theme_minimal()
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+```
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+```{r}
+ #| message: false
+ #| echo: false
+ #| include: false
+ #| label: tbl-top_neighborhood_table
+ #| tbl-cap: Top Neighborhood Table
+
+ top_hood_table <- cleaned_homicide_data %>%
+   group_by(hood_id) %>%
+   summarise(total_homicides = n()) %>%
+   arrange(desc(total_homicides)) %>%
+   head(5) %>%
+   inner_join(cleaned_homicide_data, by = "hood_id") %>%
+   select(hood_id, season, day_of_week, occurrence_year, homicide_type)
+
+ # Print the resulting table
+ print(top_hood_table)
+
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| include: false
+#| label: tbl-top _eighborhood_table2
+#| tbl-cap: Top Neighborhood Table 2
+
+top5_frequent_hood <- cleaned_homicide_data %>%
+  group_by(hood_id) %>%
+  summarise(total_occurrences = n()) %>%
+  arrange(desc(total_occurrences)) %>%
+  head(5)
+
+# Print or further analyze the top 5 frequent divisions
+top5_frequent_hood
+```
+```{r}
+# #| message: false
+# #| echo: false
+# #| label: tbl-top5_season
+# #| tbl-cap: Homicide Cases in Most Frequent Neighborhoods by Season
+# 
+# # Find the top 5 neighborhoods
+# top_neighborhoods <- cleaned_homicide_data |>
+#   group_by(hood_id) |>
+#   summarise(total_cases = n()) |>
+#   arrange(desc(total_cases)) |>
+#   top_n(5)
+# 
+# # Filter data for the top 5 neighborhoods
+# top_neighborhood_data <- cleaned_homicide_data |>
+#   filter(hood_id %in% top_neighborhoods$hood_id)
+# 
+# # Create a table with the relationship between season and the number of cases in the top 5 neighborhoods
+# season_relationship_table <- top_neighborhood_data |>
+#   group_by(hood_id, season) |>
+#   summarise(total_cases = n()) |>
+#   arrange(hood_id, desc(total_cases))
+# 
+# season_relationship_table$hood_id <- factor(season_relationship_table$hood_id,
+#                                              levels = c("73", "25", "132", "26", "2"))
+# # Transform the dataframe to the format required for printing as a table
+# season_relationship_df <- season_relationship_table |>
+#   select(hood_id, season, total_cases) |>
+#   group_by(season, hood_id) |>
+#   summarise(n = sum(total_cases)) |>
+#   pivot_wider(names_from = hood_id, values_from = n) %>%
+#   replace(is.na(.), 0) |>
+#   ungroup() |>
+#   arrange(season)
+# 
+# # Use kbl for the entire table creation and styling
+# kbl(season_relationship_df, booktabs = TRUE, linesep = "", 
+#     caption = "Homicide Cases in Most Frequent Neighborhoods by Season", 
+#     position = 'h', latex_options = "striped", font_size = 8, full_width = FALSE) |>
+#   column_spec(1, width = "2cm") |>
+#   add_header_above(c("", "Top 5 Neighborhoods" = 5), bold = TRUE)
+```
+
+```{r}
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| label: tbl-top5_season
+#| tbl-cap: Homicide Cases in Most Frequent Neighborhoods by Season
+
+# Find the top 5 neighborhoods
+top_neighborhoods <- cleaned_homicide_data |>
+  group_by(hood_id) |>
+  summarise(total_cases = n()) |>
+  arrange(desc(total_cases)) |>
+  top_n(5)
+
+# Filter data for the top 5 neighborhoods
+top_neighborhood_data <- cleaned_homicide_data |>
+  filter(hood_id %in% top_neighborhoods$hood_id)
+
+# Create a table with the relationship between season and the number of cases in the top 5 neighborhoods
+season_relationship_table <- top_neighborhood_data |>
+  group_by(hood_id, season) |>
+  summarise(total_cases = n()) |>
+  arrange(hood_id, desc(total_cases))
+
+season_relationship_table$hood_id <- factor(season_relationship_table$hood_id,
+                                             levels = c("73", "25", "132", "26", "2"))
+# Transform the dataframe to the format required for printing as a table
+ season_relationship_df <- season_relationship_table |>
+   select(hood_id, season, total_cases) |>
+   group_by(season, hood_id) |>
+   summarize(n = sum(total_cases)) |>
+   pivot_wider(names_from = hood_id, values_from = n) %>%
+   replace(is.na(.), 0) |>
+   ungroup() |>
+   arrange(season)
+
+season_relationship_df %>%
+  gt() %>%
+  tab_header(
+    title = "Homicide Cases in Most Frequent Neighborhoods by Season"
+  ) %>%
+  fmt_number(
+    columns = vars(`73`, `25`, `132`, `26`, `2`),
+    decimals = 0
+  ) %>%
+  tab_spanner(
+    label = "Top 5 Neighborhoods",
+    columns = vars(`73`, `25`, `132`, `26`, `2`)
+  ) %>%
+  fmt_number(
+    columns = vars(`73`, `25`, `132`, `26`, `2`),
+    decimals = 0
+  )
+# kbl(season_relationship_df, booktabs = TRUE, linesep = "", caption = "Homicide Cases in Most Frequent Neighborhoods by Season", position = 'h' ) |>
+#   kable_styling(latex_options = "striped", font_size = 8, full_width = FALSE) |>
+#   column_spec(1, width = "2cm") |>
+#   add_header_above(c("", "Top 5 Neighborhoods" = 5), bold = TRUE)
+
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| label: tbl-top5_dayofweek
+#| tbl-cap: homicide cases in most frequent neighborhoods by day of the week
+
+# create a table with the relationship between day of the week and the number of cases in the top 5 neighborhoods
+weekend_relationship_table <- top_neighborhood_data |>
+  group_by(hood_id, day_of_week) |>
+  summarise(total_cases = n()) |>
+  arrange(hood_id, desc(total_cases))
+
+weekend_relationship_table$hood_id <- factor(weekend_relationship_table$hood_id,
+                                             levels = c("73", "25", "132", "26", "2"))
+
+# transform the dataframe to the format required for printing as a table
+weekend_relationship_df <- weekend_relationship_table |>
+   select(hood_id, day_of_week, total_cases) |>
+   group_by(day_of_week, hood_id) |>
+   summarize(n = sum(total_cases)) |>
+   pivot_wider(names_from = hood_id, values_from = n) %>%
+   replace(is.na(.), 0) |>
+   ungroup() |>
+   arrange(day_of_week)
+
+weekend_relationship_df %>%
+  gt() %>%
+  tab_header(
+    title = "Homicide Cases in Most Frequent Neighborhoods by Season" ) %>%
+  fmt_number(
+    columns = vars(`73`, `25`, `132`, `26`, `2`),
+    decimals = 0
+  ) %>%
+  tab_spanner(
+    label = "Top 5 Neighborhoods",
+    columns = vars(`73`, `25`, `132`, `26`, `2`)
+  ) %>%
+  fmt_number(
+    columns = vars(`73`, `25`, `132`, `26`, `2`),
+    decimals = 0
+  )
+# kbl(weekend_relationship_df, booktabs = true, linesep = "", caption = "homicide cases in most frequent neighborhoods by day of the week", position = 'h') |>
+#   kable_styling(latex_options = "striped", font_size = 8, full_width = false) |>
+#   column_spec(1, width = "2cm") |>
+#   add_header_above(c("", "top 5 neighborhoods" = 5), bold = true)
+
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| label: tbl-top5_homicidetype
+#| tbl-cap: Homicide Cases in Most Frequent Neighborhoods by Type
+
+# Create a table with the relationship between homicide type and the number of cases in the top 5 neighborhoods
+type_relationship_table <- top_neighborhood_data |>
+  group_by(hood_id, homicide_type) |>
+  summarise(total_cases = n()) |>
+  arrange(hood_id, desc(total_cases))
+
+type_relationship_table$hood_id <- factor(type_relationship_table$hood_id,
+                                             levels = c("73", "25", "132", "26", "2"))
+
+# Transform the dataframe to the format required for printing as a table
+type_relationship_df <- type_relationship_table |>
+   select(hood_id, homicide_type, total_cases) |>
+   group_by(homicide_type, hood_id) |>
+   summarize(n = sum(total_cases)) |>
+   pivot_wider(names_from = hood_id, values_from = n) %>%
+   replace(is.na(.), 0) |>
+   ungroup() |>
+   arrange(homicide_type)
+
+type_relationship_df %>%
+  gt() %>%
+  tab_header(
+    title = "Homicide Cases in Most Frequent Neighborhoods by Season" ) %>%
+  fmt_number(
+    columns = vars(`73`, `25`, `132`, `26`, `2`),
+    decimals = 0
+  ) %>%
+  tab_spanner(
+    label = "Top 5 Neighborhoods",
+    columns = vars(`73`, `25`, `132`, `26`, `2`)
+  ) %>%
+  fmt_number(
+    columns = vars(`73`, `25`, `132`, `26`, `2`),
+    decimals = 0
+  )
+# 
+# kbl(type_relationship_df, booktabs = TRUE, linesep = "", caption = "Homicide Cases in Most Frequent Neighborhoods by Type", position = 'h') |>
+#   kable_styling(latex_options = "striped", font_size = 8, full_width = FALSE) |>
+#   column_spec(1, width = "2cm") |>
+#   add_header_above(c("", "Top 5 Neighborhoods" = 5), bold = TRUE)
+
+```
+
+```
+
+ggplot(cleaned_homicide_data, aes(x = division,  y = occurrence_year, fill = homicide_count)) +
+  geom_tile() + 
+  labs(title = "Correlation Between Division and Homicide Cases (Heat Map)",
+       x = "Division",
+       y = "Year",
+       fill = "Homicide Cases") +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+# 
+# Talk more about it.
+# 
+# And also planes (@fig-planes).
+(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+# 
+# ```{r}
+# #| label: fig-planes
+# #| fig-cap: Relationship between wing length and width
+# #| echo: false
+# #| warning: false
+# #| message: false
+# 
+# analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))
+# 
+# analysis_data |> 
+#   ggplot(aes(x = width, y = length)) +
+#   geom_point(alpha = 0.8) +
+#   theme_minimal() +
+#   labs(x = "Wing width (mm)",
+#        y = "Wing length (mm)")
+```
 
@@
 
-Our results are summarized in @tbl-modelresults.
+#Our results are summarized in @tbl-modelresults.
 
@@
 
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("outputs/models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
+# library(rstanarm)
+# 
+# first_model <-
+#   readRDS(file = here::here("outputs/models/first_model.rds"))
 ```
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+#In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
 
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
+#In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
 
 ## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
 

### 2024-01-23 scripts/02-data_cleaning.R

@@ -12,6 +12,7 @@ library(tidyverse)
 library(janitor)
 library(dplyr)
 library(readr)
+library(stringr)
 
 #### Clean data ####
 raw_data <- read_csv("inputs/data/raw_homicide_data.csv")
@@ -48,9 +49,9 @@ cleaned_homicide_data <-
 # find the day of the week of the homicide and determine if it is a weekend #
 cleaned_homicide_data <-
   cleaned_homicide_data |>
-  mutate(day_of_week = weekdays(occurrence_date),
-         weekend = ifelse(day_of_week %in% c("Saturday", "Sunday"), TRUE,
-                          FALSE))
+  mutate(day_of_week = weekdays(occurrence_date))
+         # weekend = ifelse(day_of_week %in% c("Saturday", "Sunday"), TRUE,
+         #                  FALSE))
                         
 # find the season when the homicide occurred #
 cleaned_homicide_data <-
@@ -62,12 +63,18 @@ cleaned_homicide_data <-
            TRUE ~ "Winter")
   )
 
-# Count the occurrence of homicides by year #
-cleaned_homicide_data <- 
-  cleaned_homicide_data |>
-  group_by(occurrence_year)|>
-  mutate(homicide_count = n())
-head(cleaned_homicide_data)
+# Extract only the date part (YYYY-MM-DD) from the date variable #
+
+cleaned_homicide_data$occurrence_date <- 
+  str_sub(cleaned_homicide_data$occurrence_date, 1, 10)
+
+
+# Count the occurrence of homicides every year #
+ cleaned_homicide_data <- 
+    cleaned_homicide_data |>
+    group_by(occurrence_year)|>
+    mutate(homicide_count = n())
+  head(cleaned_homicide_data)
 
 
 #### Save data ####

### 2024-01-23 outputs/paper/paper.qmd
fixed something

--- 2024-01-20 outputs/paper/paper.qmd
@@
 title: "Hannah Yu"
-subtitle: "My subtitle if needed"
 author: 
   - Hannah Yu
-thanks: "Code and data are available at: ."
+thanks: "Code and data are available at: "
 date: today
@@
 number-sections: true
+toc: true
 bibliography: references.bib
@@
 library(tidyverse)
-library(palmerpenguins)
 library(ggplot2)
+library(kableExtra)
+library(gt)
 
@@
 
-# Introduction
+## Introduction
 
-You can and should cross-reference sections and sub-sections.
+The homicide rate plays a pivotal role in assessing the state of violence in society.
+In 2022, Canadian police services documented 874 cases of homicide, with 2.25 homicides per 100,000 people.
+With an 8% increase in the national homicide rate, this increase signifies the fourth consecutive annual rise in the national homicide rate [@citestatscan].
+As one of Canada’s most vibrant and diverse cities, Toronto is the home of 2.93 million people.
+Toronto’s safety and stability are extremely important for all the residents of the city.
+Understanding the patterns, factors, and implications is important in ensuring Toronto’s public safety and establishing targeted interventions.
+
+This research seeks to analyze the spatial and temporal distribution of Toronto’s homicide cases.
+By investigating whether specific times or certain areas are prone to be associated with homicide incidents, we want to shed light on the dynamics of and think about addressing and preventing future prevention for homicide cases.
+
+The Police Annual Statistical Report - Homicides [@citehomicidedata] from [@ropendatatoronto] provides a record of all homicide cases in Toronto from 2004-2020 including the time of occurrence, location, and type of homicide.
+we could draw information about homicide cases and their related characteristics from this extensive dataset, and investigate Toronto’s homicide patterns across time.
+In Section 2, I introduce the dataset and some general information.
+In Section 2.1 ## General Homicide Analysis, I navigate the change in homicide case numbers across time with various characteristics and perform some analysis on the relationship between homicide cases and time variables.
+And in Section 2.2, I redo most of the analysis on the top five neighborhood with the most amount of homicide cases.
+Lastly, I discuss the implications and further discussions on these results in Section 3.
 
@@
 
-# Data {#sec-data}
+## Data {#sec-data}
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+To investigate the homicide cases in Toronto, I obtained the dataset “ Police Annual Statistical Report - Homicides [@citehomicidedata] from the Toronto Open Data Portal [@ropendatatoronto].
+Data was cleaned and analyzed using the open source statistical programming language R [@citer], with additional support from 'tidyverse' [@rtidyverse], 'dplyr' [@rdplyr], 'janitor'[@rjanitor], 'readr' [@rreadr] , 'stringr' [@rstringr], 'lubridate' [@rlubridate], 'ggplot2' [@rggplot2], 'kableExtra' [@rkableExtra], ''  Further details about data extraction and analyzation processes will be discussed in the following sections below.
+
+This dataset includes a thorough list of alongside every homicide case in Toronto from the year 2004-2020.
+We cleaned the dataset and selected variables that are relevant to our analysis of homicide cases.
+Our variable of interest includes the year the offence occurred, type of homicide (shooting, stabbing, other), date of offence, and identifier of neighborhood using Toronto’s 140 neighborhood structure.
+In addition to these variables, we also constructed three more variables: day of the week when the offence occurred (Monday-Sunday), season when the offence occurred (fall, winter, spring, summer), and count of homicide cases by year to further aid our data analysis [@tbl-cleaned_data].
+This dataset was last refreshed on January 11, 2024.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| message: false
 #| echo: false
+#| label: tbl-cleaned_data
+#| tbl-cap: Sample of Cleaned Homicide Data
+
+head (cleaned_homicide_data, 5) |>
+  kable(
+    col.names = c("Year", "Homicide Type", "Date", "Hood ID", "Day of the Week", "Season", "Homicide Count"),
+  booktabs = TRUE
+  )
+
+```
+## General Homicide Analysis
+
+We listed out the number of homicide cases each year in @tbl-homicide_count and created the summary statistics table of the homicide count per year.
+We found the year 2011 has the least cases of homicide with only 51 cases while the year 2018 with the total of 97 cases has the most number of homicide cases.
+
+```{r}
+#| message: false
+#| echo: false
+#| label: tbl-homicide_count
+#| tbl-cap: Total Homicide Cases Each Year
+
+# Assuming 'occurrence_year' is the column containing the year information
+homicide_counts_by_year <- cleaned_homicide_data |>
+  group_by(occurrence_year) |>
+  summarise(total_cases = n())
+#installed.packages("git")
+
+homicide_counts_by_year |>
+  gt() |>
+  tab_header(
+    title = "Total Homicide Cases Each Year"
+  ) |>
+  fmt_number(
+    columns = c(total_cases),
+    decimals = 0
+  )
+
+```
+Of all the homicide cases, the cases caused by shooting is significantly higher than other type of homicides.
+In the years of our examination, shooting takes up more than 600 cases of homicides while stabbing takes over more than 250, and all other types together takes up around 300 (@fig-distribution_by_type).
+
+```{r,out.width='70%', echo=FALSE}
+#| message: false
+#| echo: false
+#| label: fig-distribution_by_type
+#| fig-cap: Distribution of Homicide Types
 
@@
 
-# ggplot(penguins, aes(x = island, fill = species)) +
-#   geom_bar(alpha = 0.8) +
-#   scale_fill_manual(values = c("darkorange","purple","cyan4"),
-#                     guide = "none") +
-#   theme_minimal() +
-#   facet_wrap(~species, ncol = 1) +
-#   coord_flip()
 ```
+We then shift our attention to the homicide trend over time.
+@fig-occurrence portrays the trend of homicide cases throughout the span of almost 2 decades.
+With no obvious trend, We do not observe a clear increase or decrease of cases across time.
+However, the homicide cases are decreasing after it peaked in 2018.
 
 ```{r}
-ggplot(cleaned_homicide_data, aes(x = occurrence_year, y = count, group = 1)) +
+#| message: false
+#| echo: false
+#| label: fig-occurrence
+#| fig-cap: Homicide Occurrences Over Years
+summarized_homicide_data <- cleaned_homicide_data |>
+  group_by(occurrence_year) |>
+  summarise(homicide_count = n())
+
+ggplot(cleaned_homicide_data, aes(x = occurrence_year, y = homicide_count, group = 1)) +
   geom_line(color = "blue") +
   geom_point(color = "blue") +
-  labs(title = "Homicide Occurrences Over the Years",
+  labs(title = "Homicide Occurrences Over Years",
        x = "Year",
@@
 ```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+Following the grand picture across time, @fig-specific_occurrence breaks down the homicide cases over time into further details based on the type of homicide.
+The trend of the three types of homicide cases generally follow the overall trend in @fig-occurrence.
+Shooting always prevail over other types of homicide cases even in its lowest year.
+We conclude that there is no clear increase or decrease of homicide across time.
 
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
+#| message: false
 #| echo: false
-#| warning: false
-#| message: false
+#| label: fig-specific_occurrence
+#| fig-cap: Total Homicide Cases Over Years by Homicide Type
 
-analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))
+count_by_type <- cleaned_homicide_data |>
+  group_by(occurrence_year, homicide_type) |>
+  summarise(total_cases = n(), .groups = 'drop_last')
 
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
+ggplot(count_by_type,
+       aes(x = occurrence_year, y = total_cases, color = homicide_type)) +
+  geom_line() + 
+  labs(title = "Total Homicide Cases Over Years by Homicide Type",
+       x = "Year",
+       y = "Total Cases",
+       color = "Homicide Type") +
+  theme_minimal()
+
 ```
 
-Talk way more about it.
-
-# Model
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
+We want to determine whether if there are a specific time in the year when homicides are more likely to occurr.
+To investigate our interest, we created @fig-case_by_week that displays the homicide counts by every day of the week.
+Sunday with a total of 200 cases is the day with the most cases following by Saturday and Friday with 197 and 175 respectively.
+Thursday is the day with the least cases with 124 cases.
+We conclude that homicides are slightly more likely to occur in weekends.
 
 ```{r}
+#| message: false
 #| echo: false
-#| eval: true
-#| warning: false
-#| message: false
+#| include: false
+#| label: tbl-casebyweek
+#| tbl-cap: Case_by_week
 
-library(rstanarm)
+homicide_counts_by_day <- cleaned_homicide_data %>%
+  group_by(day_of_week) %>%
+  summarise(total_cases = n())
 
-first_model <-
-  readRDS(file = here::here("outputs/models/first_model.rds"))
+print(homicide_counts_by_day)
+
 ```
@@
 ```{r}
+#| message: false
 #| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
+#| label: fig-case_by_week
+#| fig-cap: Distribution of Homicide Cases by Day of the Week
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
+cleaned_homicide_data$day_of_week <- factor(cleaned_homicide_data$day_of_week, 
+                                             levels = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
+
+ggplot(cleaned_homicide_data, aes(x = day_of_week, fill = day_of_week)) +
+  geom_bar() +
+  labs(title = "Distribution of Homicide Cases by Day of the Week",
+       x = "Day of the Week",
+       y = "Number of Homicides") +
+  scale_x_discrete(labels = c("Monday" = "Mon.", "Tuesday" = "Tue.", "Wednesday" = "Wed.", 
+                              "Thursday" = "Thu.", "Friday" = "Fri.", "Saturday" = "Sat.", "Sunday" = "Sun.")) +
+  theme_minimal()
+
 ```
 
-# Discussion
+We perform similar analysis with the count of homicide cases by season and found that the seasonal data are distributed somewhat evenly with fall having slightly more cases than other weather but the difference is to small and not worth noting @fig-case_by_season.
+```{r}
+#| message: false
+#| echo: false
+#| label: fig-case_by_season
+#| fig-cap: Distribution of Homicide Cases by Season
 
-## First discussion point {#sec-first-point}
+cleaned_homicide_data$season <- factor(cleaned_homicide_data$season, 
+                                             levels = c("Fall", "Winter", "Spring", "Summer"))
+ggplot(cleaned_homicide_data, aes(x = factor(season), fill = season)) +
+  geom_bar() +
+  labs(title = "Distribution of Homicide Cases by Season",
+       x = "Season",
+       y = "Number of Homicides") +
+  theme_minimal()
+```
+To investigate the relationship between geographical location and homicide cases, we constructed @fig-neighborhoods_distribution.
+This bar plot illustrates the distribution of homicide cases in Toronto's 140 neighborhoods.
+This visualization enables the identification of patterns or areas of concerns in the city.
+While most neighborhoods have fewer than 10 cases, we notice that several neighborhoods stand out with significantly higher homicide incidences compared to the rest.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+```{r}
+#| message: false
+#| echo: false
+#| label: fig-neighborhoods_distribution
+#| fig-cap: Distribution of Homicide in Neighborhoods
 
-## Second discussion point
+nei<- cleaned_homicide_data %>%
+  group_by(hood_id) %>%
+  summarise(total_cases = n())
 
-## Third discussion point
+ggplot(cleaned_homicide_data, aes(x = hood_id)) +
+  geom_bar(fill = "skyblue", color = "black") +
+  labs(title = "Distribution of Homicide Cases in 140 Neighborhoods",
+       x = "Neighborhood Number",
+       y = "Number of Homicide Cases") +
+  theme_minimal()
 
-## Weaknesses and next steps
+```
+## Homicide Analysis in Top Five Neighborhood
 
-Weaknesses and next steps should also be included.
+Having observed the obvious significantly higher cases of homicide in several neighborhoods, I selected the five neighborhoods with the highest amount of homicide cases and performed the following analysis that focus on these five neighborhoods.
+The ids of the five neighborhoods are 73, 25, 132, 26, 2 respectively with neighborhood 73 having the highest number of cases of 31 and neighborhood 2 with 28 cases.
+Through these analysis, we could understand if there is specific patterns or trends in the occurrence of homicide incidences in these top five neighborhoods and genearlize the conclusion in larger areas.
+
+In order to observe the relationship between relating factors and homicide cases, I investigate the correlation of the occurrence of each case with each day of the week, each season, and type of homicide (@tbl-top5_dayofweek, @tbl-top5_season, @tbl-top5_homicidetype).
+We do not observe a specific trend in the relation between homicide cases and day of the week.
+Just like our previous analysis on @fig-case_by_season, the correlation between the number of homicide cases and with a specific day of the week or the weekends are also not noticeable.
+We could also attribute this result to the fact that our sample size is not big enough to observe a difference as there are only limited amount of homicide cases in the city.
+However, we do notice that just like the previous, the amount of homicide caused by shooting is higher than any other methods in each of the top five neighborhoods.
+This is a confirmation that shooting is the main type of homicide in Toronto.
+```{r}
+#| message: false
+#| include: false
+#| label: tbl-top _neighborhood_table2
+#| tbl-cap: Top Neighborhood Table 2
+
+top5_frequent_hood <- cleaned_homicide_data %>%
+  group_by(hood_id) %>%
+  summarise(total_occurrences = n()) %>%
+  arrange(desc(total_occurrences)) %>%
+  head(5)
+
+# Print or further analyze the top 5 frequent divisions
+print(top5_frequent_hood)
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| label: tbl-top5_season
+#| tbl-cap: Homicide Cases in Most Frequent Neighborhoods by Season
+
+# Find the top 5 neighborhoods
+top_neighborhoods <- cleaned_homicide_data |>
+  group_by(hood_id) |>
+  summarise(total_cases = n()) |>
+  arrange(desc(total_cases)) |>
+  top_n(5)
+
+# Filter data for the top 5 neighborhoods
+top_neighborhood_data <- cleaned_homicide_data |>
+  filter(hood_id %in% top_neighborhoods$hood_id)
+
+# Create a table with the relationship between season and the number of cases in the top 5 neighborhoods
+season_relationship_table <- top_neighborhood_data |>
+  group_by(hood_id, season) |>
+  summarise(total_cases = n()) |>
+  arrange(hood_id, desc(total_cases))
+
+season_relationship_table$hood_id <- factor(season_relationship_table$hood_id,
+                                             levels = c("73", "25", "132", "26", "2"))
+
+# Transform the dataframe to the format required for printing as a table
+ season_relationship_df <- season_relationship_table |>
+   select(hood_id, season, total_cases) |>
+   group_by(season, hood_id) |>
+   summarize(n = sum(total_cases)) |>
+   pivot_wider(names_from = hood_id, values_from = n) %>%
+   replace(is.na(.), 0) |>
+   ungroup() |>
+   arrange(season)
+
+season_relationship_df |>
+  gt() |>
+  tab_header(
+    title = "Homicide Cases in Most Frequent Neighborhoods by Season"
+  )  |>
+  fmt_number(
+    columns = c(`73`, `25`, `132`, `26`, `2`),
+    decimals = 0
+  )  |>
+  tab_spanner(
+    label = "Top 5 Neighborhoods",
+    columns = c(`73`, `25`, `132`, `26`, `2`)
+  ) |>
+  fmt_number(
+    columns = c(`73`, `25`, `132`, `26`, `2`),
+    decimals = 0
+  )
+
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| label: tbl-top5_dayofweek
+#| tbl-cap: homicide cases in most frequent neighborhoods by day of the week
+
+# create a table with the relationship between day of the week and the number of cases in the top 5 neighborhoods
+weekend_relationship_table <- top_neighborhood_data |>
+  group_by(hood_id, day_of_week) |>
+  summarise(total_cases = n()) |>
+  arrange(hood_id, desc(total_cases))
+
+weekend_relationship_table$hood_id <- factor(weekend_relationship_table$hood_id,
+                                             levels = c("73", "25", "132", "26", "2"))
+
+# transform the dataframe to the format required for printing as a table
+weekend_relationship_df <- weekend_relationship_table |>
+   select(hood_id, day_of_week, total_cases) |>
+   group_by(day_of_week, hood_id) |>
+   summarize(n = sum(total_cases)) |>
+   pivot_wider(names_from = hood_id, values_from = n) %>%
+   replace(is.na(.), 0) |>
+   ungroup() |>
+   arrange(day_of_week)
+
+weekend_relationship_df |>
+  gt()|>
+  tab_header(
+    title = "Homicide Cases in Most Frequent Neighborhoods by Day of the Week" ) |>
+  fmt_number(
+    columns = c(`73`, `25`, `132`, `26`, `2`),
+    decimals = 0
+  ) |>
+  tab_spanner(
+    label = "Top 5 Neighborhoods",
+    columns = c(`73`, `25`, `132`, `26`, `2`)
+  )  |>
+  fmt_number(
+    columns = c(`73`, `25`, `132`, `26`, `2`),
+    decimals = 0
+  )
+
+```
+
+```{r}
+#| message: false
+#| echo: false
+#| label: tbl-top5_homicidetype
+#| tbl-cap: Homicide Cases in Most Frequent Neighborhoods by Type
+
+# Create a table with the relationship between homicide type and the number of cases in the top 5 neighborhoods
+type_relationship_table <- top_neighborhood_data |>
+  group_by(hood_id, homicide_type) |>
+  summarise(total_cases = n()) |>
+  arrange(hood_id, desc(total_cases))
+
+type_relationship_table$hood_id <- factor(type_relationship_table$hood_id,
+                                             levels = c("73", "25", "132", "26", "2"))
+
+# Transform the dataframe to the format required for printing as a table
+type_relationship_df <- type_relationship_table |>
+   select(hood_id, homicide_type, total_cases) |>
+   group_by(homicide_type, hood_id) |>
+   summarize(n = sum(total_cases)) |>
+   pivot_wider(names_from = hood_id, values_from = n) %>%
+   replace(is.na(.), 0) |>
+   ungroup() |>
+   arrange(homicide_type)
+
+type_relationship_df |>
+  gt()  |>
+  tab_header(
+    title = "Homicide Types in Most Frequent Neighborhoods" )  |>
+  fmt_number(
+    columns = c(`73`, `25`, `132`, `26`, `2`),
+    decimals = 0
+  ) %>%
+  tab_spanner(
+    label = "Top 5 Neighborhoods",
+    columns = c(`73`, `25`, `132`, `26`, `2`)
+  ) %>%
+  fmt_number(
+    columns = c(`73`, `25`, `132`, `26`, `2`),
+    decimals = 0
+  )
+
+```
+`
+
+## Discussion
+
+The result of our general analysis of homicide cases by season and by each day of the week throughout Toronto generally align with our specific focus into the top five highest homicide cases neighborhoods.
+We cannot obverse any significant correlation between homicidece occurrences and particular timings.
+Homicides may have a slightly higher chance in occurring on a weekend than weekdays and are not more likely to occur in a particular season.
+
+However, one characteristics that stand out throughout the entirety of our analysis is that how shooting is always the highest type of homicide.
+It started as the highest cause of homicide in 2004 and remained until 2020.
+We could conclude that is the highest cause of homicide across time.
+
+We do not observe a general increase or decrease of homicide cases throughout time.
+However, afte the number of homicide cases peaked in 2018, the cases are experiencing a decline.
+Cases caused by shooting are also in decline during the same period.
+This decline may be associated with how Toronto gun violence has dropped by 30% over the past five years (@dropingun).
+
+Our analysis would be more comprehensive if there are more detailed data about more relevant characteristics about homicides and more cases of homicide.
+But due to the scope of our dataset and the nature of homicide occurrences being generally rare compared to other research topics, it is very hard to give out a firm statement such as a certain factor causes more homicide incidences or homicide incidences are more likely to happen on a certain dates.
 
@@
 
-\appendix
-
-# Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
-
-\newpage
-
-# References
+## References

### 2024-01-23 scripts/00-simulate_data.R

@@ -14,6 +14,8 @@ library(knitr)
 library(janitor)
 library(lubridate)
 library(opendatatoronto)
+library(ggplot2)
+
 
 #### Simulate data ####
 
@@ -50,7 +52,6 @@ print(simulated_data)
 
 
 #Bar Plot for Homicide Type
-library(ggplot2)
 
 
 ggplot(simulated_data, aes(x = homicide_type, fill = homicide_type)) +

### 2024-01-23 scripts/02-data_cleaning.R

@@ -39,7 +39,7 @@ cleaned_homicide_data <-
   cleaned_homicide_data |>
   select(
     occurrence_year, 
-    division,
+#    division,
     homicide_type,
     occurrence_date,
     hood_id

### 2024-01-23 outputs/paper/paper.qmd
deleted some files

@@
   - Hannah Yu
-thanks: "Code and data are available at: "
+thanks: "Code and data are available at: https://github.com/hannahyu07/Toronto-Homicide-Analysis"
 date: today
@@
 
-## Introduction
+# Introduction
 
@@
 we could draw information about homicide cases and their related characteristics from this extensive dataset, and investigate Toronto’s homicide patterns across time.
-In Section 2, I introduce the dataset and some general information.
-In Section 2.1 ## General Homicide Analysis, I navigate the change in homicide case numbers across time with various characteristics and perform some analysis on the relationship between homicide cases and time variables.
-And in Section 2.2, I redo most of the analysis on the top five neighborhood with the most amount of homicide cases.
-Lastly, I discuss the implications and further discussions on these results in Section 3.
+In @sec-data, I introduce the dataset and some general information.
+In @subsec1-data, I navigate the change in homicide case numbers across time with various characteristics and perform some analysis on the relationship between homicide cases and time variables.
+And in @subsec2-data, I redo most of the analysis on the top five neighborhood with the most amount of homicide cases.
+Lastly, I discuss the implications and further discussions on these results in @sec-discussion.
 
-The remainder of this paper is structured as follows. @sec-data....
-
-## Data {#sec-data}
+# Data {#sec-data}
 
 To investigate the homicide cases in Toronto, I obtained the dataset “ Police Annual Statistical Report - Homicides [@citehomicidedata] from the Toronto Open Data Portal [@ropendatatoronto].
-Data was cleaned and analyzed using the open source statistical programming language R [@citer], with additional support from 'tidyverse' [@rtidyverse], 'dplyr' [@rdplyr], 'janitor'[@rjanitor], 'readr' [@rreadr] , 'stringr' [@rstringr], 'lubridate' [@rlubridate], 'ggplot2' [@rggplot2], 'kableExtra' [@rkableExtra], ''  Further details about data extraction and analyzation processes will be discussed in the following sections below.
+Data was cleaned and analyzed using the open source statistical programming language R [@citer], with additional support from 'tidyverse' [@rtidyverse], 'dplyr' [@rdplyr], 'janitor'[@rjanitor], 'readr' [@rreadr] , 'stringr' [@rstringr], 'lubridate' [@rlubridate], 'ggplot2' [@rggplot2], 'kableExtra' [@rkableExtra], 'gt'[@rgt].
+Further details about data extraction and analyzation processes will be discussed in the following sections below.
 
@@
 ```
-## General Homicide Analysis
+## General Homicide Analysis {#subsec1-data}
 
@@
 ```
-## Homicide Analysis in Top Five Neighborhood
+## Homicide Analysis in Top Five Neighborhood {#subsec2-data}
 
@@
 
-## Discussion
+# Discussion {#sec-discussion}
 

### 2024-01-23 scripts/00-simulate_data.R

@@ -1,11 +1,11 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
+# Purpose: Simulates homicide data
 # Author: Hannah Yu [...UPDATE THIS...]
 # Date: 17 January 2024 [...UPDATE THIS...]
-# Contact: realhannah.yu@mail.utoronto.ca [...UPDATE THIS...]
+# Contact: realhannah.yu@mail.utoronto.ca 
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: None
+
 
 
 #### Workspace setup ####
@@ -35,14 +35,14 @@ num_obs <- 1398
 
 simulated_data <- 
   tibble(
-    year = sample(2004:2023, num_obs, replace = TRUE),
+    year = sample(2004:2020, num_obs, replace = TRUE),
     month = sample(1:12, num_obs, replace = TRUE),
     day = sample(1:31, num_obs, replace = TRUE),
     weekday = sample(c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday",
                        "Friday", "Saturday"), num_obs, replace = TRUE),
     homicide_type = sample(c("Shooting", "Stabbing", "Other"),  num_obs,
                            replace = TRUE),
-    neighborhood_number = sample(1:158, num_obs, replace = TRUE),
+    neighborhood_number = sample(1:140, num_obs, replace = TRUE),
 )
 
 
@@ -70,17 +70,17 @@ ggplot(simulated_data, aes(x = neighborhood_number)) +
 # Check the year of homicide occurrence is between 2004-2023
 
 simulated_data$year |> min() == 2004
-simulated_data$year |> max() == 2023
+simulated_data$year |> max() == 2020
 
 # Check the day of homicide occurrence is between 1-31
 
 simulated_data$day |> min() == 1
 simulated_data$day |> max() == 31
 
-# Check the neighborhood number is between 1-158
+# Check the neighborhood number is between 1-140
 
 simulated_data$neighborhood_number |> min() == 1
-simulated_data$neighborhood_number |> max() == 158
+simulated_data$neighborhood_number |> max() == 140
 
 # Check homicide type is character 
 

### 2024-01-23 scripts/01-download_data.R

@@ -1,11 +1,10 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Hannah Yu [...UPDATE THIS...]
-# Date: 17 January 2024 [...UPDATE THIS...]
-# Contact: realhannah.yu@mail.utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from OpenData Toronto portal
+# Author: Hannah Yu 
+# Date: 17 January 2024  
+# Contact: realhannah.yu@mail.utoronto.ca 
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: None
 
 
 #### Workspace setup ####

### 2024-01-23 scripts/02-data_cleaning.R

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Cleans the raw homicide data by and selecting relevant columns. 
+# Purpose: Cleans the raw homicide data and select relevant columns. 
 # Author: Hannah Yu
 # Date: 19 January 2023  
 # Contact: realhannah.yu@mail.utoronto.ca 
@@ -86,32 +86,4 @@ write_csv(
 write_csv(cleaned_homicide_data, "outputs/data/analysis_data.csv")
 
 
-# cleaned_data <-
-#   raw_data |>
-#   janitor::clean_names() |>
-#   select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-#   filter(wing_width_mm != "caw") |>
-#   mutate(
-#     flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-#                                    "1.35",
-#                                    flying_time_sec_first_timer)
-#   ) |>
-#   mutate(wing_width_mm = if_else(wing_width_mm == "490",
-#                                  "49",
-#                                  wing_width_mm)) |>
-#   mutate(wing_width_mm = if_else(wing_width_mm == "6",
-#                                  "60",
-#                                  wing_width_mm)) |>
-#   mutate(
-#     wing_width_mm = as.numeric(wing_width_mm),
-#     wing_length_mm = as.numeric(wing_length_mm),
-#     flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-#   ) |>
-#   rename(flying_time = flying_time_sec_first_timer,
-#          width = wing_width_mm,
-#          length = wing_length_mm
-#          ) |> 
-#   tidyr::drop_na()
-# 
-# #### Save data ####
-# write_csv(cleaned_data, "outputs/data/analysis_data.csv")
+

### 2024-01-23 scripts/03-test_data.R

@@ -1,15 +1,28 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests cleaned data
+# Author: Hannah Yu
+# Date: 23 January 2024 
+# Contact: realhannah.yu@mail.utoronto.ca  
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
+# Pre-requisites: 02-data_cleaning.R
+ 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+
 
 #### Test data ####
+# Check the year of homicide occurrence is between 2004-2020
+
+cleaned_homicide_data$occurrence_year |> min() == 2004
+cleaned_homicide_data$occurrence_year |> max()== 2020
+
+
+# Check the neighborhood number is between 1-140
+
+cleaned_homicide_data$hood_id |> min() == 1
+cleaned_homicide_data$hood_id |> max() == 140
+
+# Check homicide type is character 
+
+cleaned_homicide_data$homicide_type |> class() == "character"

### 2024-01-23 outputs/paper/paper.qmd
edited paragraphs and added llm

@@
 ---
-title: "Hannah Yu"
+title: "No evidence of Homicide Patterns: A Comprehensive Analysis of Homicide Cases and Characteristics in Toronto (2004-2020)"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "One crucial factor in measuring a city’s well-being is its homicide rates and its associating characteristics.
+Using data from OpenDataToronto, this paper aims to investigate all recorded homicide cases in Toronto from 2004-2020.
+By exploring data encompassing all Toronto neighbourhoods and focusing those with the highest incidence of homicide cases, I found that shooting is the predominant form of homicide.
+This paper reveals that homicide occurrences in Toronto, in general, do not follow a noticeable temporal pattern, as there is no correlation between the frequency of homicides and temporal variables."
 format: pdf
@@
 library(ggplot2)
-library(kableExtra)
 library(gt)
+library(knitr)
 
@@
 The homicide rate plays a pivotal role in assessing the state of violence in society.
-In 2022, Canadian police services documented 874 cases of homicide, with 2.25 homicides per 100,000 people.
+In 2022, Canadian police services documented 874 cases of homicide, with 2.25 homicides per 100,000 people [@citestatscan].
 With an 8% increase in the national homicide rate, this increase signifies the fourth consecutive annual rise in the national homicide rate [@citestatscan].
@@
 Toronto’s safety and stability are extremely important for all the residents of the city.
-Understanding the patterns, factors, and implications is important in ensuring Toronto’s public safety and establishing targeted interventions.
+Understanding the patterns, factors, and implications of homicide cases is important in ensuring Toronto’s public safety and establishing targeted interventions.
 
@@
 
-The Police Annual Statistical Report - Homicides [@citehomicidedata] from [@ropendatatoronto] provides a record of all homicide cases in Toronto from 2004-2020 including the time of occurrence, location, and type of homicide.
-we could draw information about homicide cases and their related characteristics from this extensive dataset, and investigate Toronto’s homicide patterns across time.
-In @sec-data, I introduce the dataset and some general information.
-In @subsec1-data, I navigate the change in homicide case numbers across time with various characteristics and perform some analysis on the relationship between homicide cases and time variables.
-And in @subsec2-data, I redo most of the analysis on the top five neighborhood with the most amount of homicide cases.
-Lastly, I discuss the implications and further discussions on these results in @sec-discussion.
+The Police Annual Statistical Report - Homicides [@citehomicidedata] from [@ropendatatoronto] provides a record of all homicide cases in Toronto from 2004-2020 including details such as the time of occurrence, location, and type of homicide.
+I could draw information about homicide cases and their related features from this extensive dataset and investigate Toronto’s homicide patterns across time.
+In @sec-data, I introduce the dataset along with general information.
+In its subsection, I navigate the change in homicide case numbers over time considering various characteristics and analyzing the relationship between homicide cases and temporal variables.
+And I refine most of the analysis specifically on the top five neighbourhoods with the highest number of homicide cases.
+Lastly, I delve into the implications and further discussions on these results in @sec-discussion.
 
@@
 To investigate the homicide cases in Toronto, I obtained the dataset “ Police Annual Statistical Report - Homicides [@citehomicidedata] from the Toronto Open Data Portal [@ropendatatoronto].
-Data was cleaned and analyzed using the open source statistical programming language R [@citer], with additional support from 'tidyverse' [@rtidyverse], 'dplyr' [@rdplyr], 'janitor'[@rjanitor], 'readr' [@rreadr] , 'stringr' [@rstringr], 'lubridate' [@rlubridate], 'ggplot2' [@rggplot2], 'kableExtra' [@rkableExtra], 'gt'[@rgt].
+Data was cleaned and analyzed using the open source statistical programming language R [@citer], with additional support from 'tidyverse' [@rtidyverse], 'dplyr' [@rdplyr], 'janitor'[@rjanitor], 'readr' [@rreadr], 'stringr' [@rstringr], 'lubridate' [@rlubridate], 'ggplot2' [@rggplot2], 'knitr' [@rknitr], 'gt'[@rgt].
 Further details about data extraction and analyzation processes will be discussed in the following sections below.
@@
 This dataset includes a thorough list of alongside every homicide case in Toronto from the year 2004-2020.
-We cleaned the dataset and selected variables that are relevant to our analysis of homicide cases.
-Our variable of interest includes the year the offence occurred, type of homicide (shooting, stabbing, other), date of offence, and identifier of neighborhood using Toronto’s 140 neighborhood structure.
-In addition to these variables, we also constructed three more variables: day of the week when the offence occurred (Monday-Sunday), season when the offence occurred (fall, winter, spring, summer), and count of homicide cases by year to further aid our data analysis [@tbl-cleaned_data].
+I cleaned the dataset and selected variables that are relevant to my analysis of homicide cases.
+My variables of interest include the year the offence occurred, type of homicide (shooting, stabbing, other), date of offence, and identifier of neighbourhood using Toronto’s 140 neighbourhood structure.
+In addition to these variables, I also constructed three more variables: day of the week when the offence occurred (Monday-Sunday), season when the offence occurred (fall, winter, spring, summer), and count of homicide cases by year to further aid our data analysis [@tbl-cleaned_data].
 This dataset was last refreshed on January 11, 2024.
@@
 
-We listed out the number of homicide cases each year in @tbl-homicide_count and created the summary statistics table of the homicide count per year.
-We found the year 2011 has the least cases of homicide with only 51 cases while the year 2018 with the total of 97 cases has the most number of homicide cases.
+I listed out the number of homicide cases each year in @tbl-homicide_count and created the summary statistics table detailing the homicide count per year.
+I found the year 2011 has the lowest cases of homicide, totaling only 51, while 2018 with the total of 97 has the most number of homicide cases.
 
@@
 ```
-Of all the homicide cases, the cases caused by shooting is significantly higher than other type of homicides.
-In the years of our examination, shooting takes up more than 600 cases of homicides while stabbing takes over more than 250, and all other types together takes up around 300 (@fig-distribution_by_type).
+Of all the homicide cases analyzed, the cases related to shooting is significantly higher than other type of homicides.
+Over the years of under examination, shooting incidents accounted for more than 600 cases of homicides, surpassing stabbing, which accounted for over 250 cases, and all other types combined which takes up around 300 cases (@fig-distribution_by_type).
 
@@
 ```
-We then shift our attention to the homicide trend over time.
-@fig-occurrence portrays the trend of homicide cases throughout the span of almost 2 decades.
-With no obvious trend, We do not observe a clear increase or decrease of cases across time.
-However, the homicide cases are decreasing after it peaked in 2018.
+We then redirect our attention to the temporal trend of homicide cases.
+@fig-occurrence portrays the trend of homicide cases throughout the span of nearly 2 decades.
+With no obvious trend evident, I do not observe a consistent increase or decrease in cases over time.
+However, it is noteworthy that the number homicide cases are decreasing after it peaked in 2018.
 
@@
 ```
-Following the grand picture across time, @fig-specific_occurrence breaks down the homicide cases over time into further details based on the type of homicide.
-The trend of the three types of homicide cases generally follow the overall trend in @fig-occurrence.
-Shooting always prevail over other types of homicide cases even in its lowest year.
-We conclude that there is no clear increase or decrease of homicide across time.
+Following our examination over time, @fig-specific_occurrence breaks down the homicide cases over time into further details based on the type of homicide.
+The trends of the three distinct types of homicide cases broadly follow the overall pattern in @fig-occurrence.
+Shooting cases always prevail over other types of homicide cases even during their lowest year.
+We conclude that there is no evident of clear increase or decrease in homicides across time.
 
@@
 
-We want to determine whether if there are a specific time in the year when homicides are more likely to occurr.
-To investigate our interest, we created @fig-case_by_week that displays the homicide counts by every day of the week.
-Sunday with a total of 200 cases is the day with the most cases following by Saturday and Friday with 197 and 175 respectively.
-Thursday is the day with the least cases with 124 cases.
-We conclude that homicides are slightly more likely to occur in weekends.
+We want to identify whether if there are specific times in the year when homicides are more likely to occur.
+To explore this topic, we generated @fig-case_by_week that displays the homicide counts by each day of the week.
+Sunday with a total count of 200 cases is the day with the most cases followed by Saturday and Friday with 197 and 175 respectively.
+On the other hand, Thursday has the lowest count, with only 124 cases.
+This analyzation suggests that homicides are slightly more likely to occur on weekends.
 
@@
 
-We perform similar analysis with the count of homicide cases by season and found that the seasonal data are distributed somewhat evenly with fall having slightly more cases than other weather but the difference is to small and not worth noting @fig-case_by_season.
+We performed a parallel analysis on the count of homicide cases by season and found a relatively even distribution with fall exhibiting slightly more cases than other seasons, but the difference is marginal @fig-case_by_season.
+
 ```{r}
@@
 ```
-To investigate the relationship between geographical location and homicide cases, we constructed @fig-neighborhoods_distribution.
-This bar plot illustrates the distribution of homicide cases in Toronto's 140 neighborhoods.
+To investigate the relationship between geographical location and homicide cases, we constructed @fig-neighbourhoods_distribution.
+This bar plot illustrates the distribution of homicide cases in Toronto's 140 neighbourhoods.
 This visualization enables the identification of patterns or areas of concerns in the city.
-While most neighborhoods have fewer than 10 cases, we notice that several neighborhoods stand out with significantly higher homicide incidences compared to the rest.
+While most neighbourhoods report fewer than 10 cases, it is noteworthy that several neighbourhoods stand out with significantly higher homicide incidences compared to the rest.
 
@@
 #| echo: false
-#| label: fig-neighborhoods_distribution
-#| fig-cap: Distribution of Homicide in Neighborhoods
+#| label: fig-neighbourhoods_distribution
+#| fig-cap: Distribution of Homicide in neighbourhoods
 
@@
   geom_bar(fill = "skyblue", color = "black") +
-  labs(title = "Distribution of Homicide Cases in 140 Neighborhoods",
-       x = "Neighborhood Number",
+  labs(title = "Distribution of Homicide Cases in 140 Neighbourhoods",
+       x = "Neighbourhood Number",
        y = "Number of Homicide Cases") +
@@
 ```
-## Homicide Analysis in Top Five Neighborhood {#subsec2-data}
+## Homicide Analysis in Top Five Neighbourhood {#subsec2-data}
 
-Having observed the obvious significantly higher cases of homicide in several neighborhoods, I selected the five neighborhoods with the highest amount of homicide cases and performed the following analysis that focus on these five neighborhoods.
-The ids of the five neighborhoods are 73, 25, 132, 26, 2 respectively with neighborhood 73 having the highest number of cases of 31 and neighborhood 2 with 28 cases.
-Through these analysis, we could understand if there is specific patterns or trends in the occurrence of homicide incidences in these top five neighborhoods and genearlize the conclusion in larger areas.
+Having observed the significantly higher cases of homicide in specific neighbourhoods, I selected the five neighbourhoods with the highest number of homicide cases-neighbourhoods 73, 25, 132, 26, and 2 with neighbourhood 73 having the highest number of cases of 31 and neighbourhood 2 with 28 cases.
+Through targeted analysis, we can discern if there is specific patterns or trends in the occurrence of homicide incidences in these neighbourhoods and generalize the conclusion in broader contexts.
 
-In order to observe the relationship between relating factors and homicide cases, I investigate the correlation of the occurrence of each case with each day of the week, each season, and type of homicide (@tbl-top5_dayofweek, @tbl-top5_season, @tbl-top5_homicidetype).
-We do not observe a specific trend in the relation between homicide cases and day of the week.
-Just like our previous analysis on @fig-case_by_season, the correlation between the number of homicide cases and with a specific day of the week or the weekends are also not noticeable.
-We could also attribute this result to the fact that our sample size is not big enough to observe a difference as there are only limited amount of homicide cases in the city.
-However, we do notice that just like the previous, the amount of homicide caused by shooting is higher than any other methods in each of the top five neighborhoods.
-This is a confirmation that shooting is the main type of homicide in Toronto.
+In order to observe the relationship between various relating factors and homicide cases, I examined the correlation of the occurrence of each case with each day of the week, season, and type of homicide (@tbl-top5_dayofweek, @tbl-top5_season, @tbl-top5_homicidetype).
+Interestingly, we do not observe a specific trend in the relation between homicide cases and day of the week.
+Similar to our earlier analysis on @fig-case_by_season, the correlation between the number of homicide cases and with a specific day of the week or the weekends are also not noticeable.
+We could attribute this result to limited sample size, given the relatively small number of homicide cases in the city compared to other crimes.
+However, we do notice that just like our earlier results, the amount of homicide caused by shooting is consistently higher than any other methods.
+This is reaffirms that shooting is the predominant type of homicide in Toronto.
+
 ```{r}
@@
 #| include: false
-#| label: tbl-top _neighborhood_table2
-#| tbl-cap: Top Neighborhood Table 2
+#| label: tbl-top _neighbourhood_table2
+#| tbl-cap: Top neighbourhood Table 2
 
@@
 #| label: tbl-top5_season
-#| tbl-cap: Homicide Cases in Most Frequent Neighborhoods by Season
+#| tbl-cap: Homicide Cases in Most Frequent neighbourhoods by Season
 
-# Find the top 5 neighborhoods
-top_neighborhoods <- cleaned_homicide_data |>
+# Find the top 5 neighbourhoods
+top_neighbourhoods <- cleaned_homicide_data |>
   group_by(hood_id) |>
@@
 
-# Filter data for the top 5 neighborhoods
-top_neighborhood_data <- cleaned_homicide_data |>
-  filter(hood_id %in% top_neighborhoods$hood_id)
+# Filter data for the top 5 neighbourhoods
+top_neighbourhood_data <- cleaned_homicide_data |>
+  filter(hood_id %in% top_neighbourhoods$hood_id)
 
-# Create a table with the relationship between season and the number of cases in the top 5 neighborhoods
-season_relationship_table <- top_neighborhood_data |>
+# Create a table with the relationship between season and the number of cases in the top 5 neighbourhoods
+season_relationship_table <- top_neighbourhood_data |>
   group_by(hood_id, season) |>
@@
   tab_header(
-    title = "Homicide Cases in Most Frequent Neighborhoods by Season"
+    title = "Homicide Cases in Most Frequent Neighbourhoods by Season"
   )  |>
@@
   tab_spanner(
-    label = "Top 5 Neighborhoods",
+    label = "Top 5 Neighbourhoods",
     columns = c(`73`, `25`, `132`, `26`, `2`)
@@
 #| label: tbl-top5_dayofweek
-#| tbl-cap: homicide cases in most frequent neighborhoods by day of the week
+#| tbl-cap: Homicide Cases in Most Frequent Neighbourhoods by Day of the Week
 
-# create a table with the relationship between day of the week and the number of cases in the top 5 neighborhoods
-weekend_relationship_table <- top_neighborhood_data |>
+# create a table with the relationship between day of the week and the number of cases in the top 5 neighbourhoods
+weekend_relationship_table <- top_neighbourhood_data |>
   group_by(hood_id, day_of_week) |>
@@
   tab_header(
-    title = "Homicide Cases in Most Frequent Neighborhoods by Day of the Week" ) |>
+    title = "Homicide Cases in Most Frequent Neighbourhoods by Day of the Week" ) |>
   fmt_number(
@@
 #| label: tbl-top5_homicidetype
-#| tbl-cap: Homicide Cases in Most Frequent Neighborhoods by Type
+#| tbl-cap: Homicide Cases in Most Frequent Neighbourhoods by Type
 
-# Create a table with the relationship between homicide type and the number of cases in the top 5 neighborhoods
-type_relationship_table <- top_neighborhood_data |>
+# Create a table with the relationship between homicide type and the number of cases in the top 5 neighbourhoods
+type_relationship_table <- top_neighbourhood_data |>
   group_by(hood_id, homicide_type) |>
@@
   tab_header(
-    title = "Homicide Types in Most Frequent Neighborhoods" )  |>
+    title = "Homicide Types in Most Frequent Neighbourhoods" )  |>
   fmt_number(
@@
 
-The result of our general analysis of homicide cases by season and by each day of the week throughout Toronto generally align with our specific focus into the top five highest homicide cases neighborhoods.
-We cannot obverse any significant correlation between homicidece occurrences and particular timings.
-Homicides may have a slightly higher chance in occurring on a weekend than weekdays and are not more likely to occur in a particular season.
+The outcome of our overall analysis of homicide cases by season and by each day of the week throughout Toronto generally align with our specific focus into the top five neighbourhoods with the highest homicide cases.
+We discern no significant correlation between homicidece occurrences and particular timings.
+Homicides may have a slightly higher likelihood in occurring on weekends compared to weekdays but there is no distinct pattern of homicide cases in a certain season.
 
-However, one characteristics that stand out throughout the entirety of our analysis is that how shooting is always the highest type of homicide.
-It started as the highest cause of homicide in 2004 and remained until 2020.
-We could conclude that is the highest cause of homicide across time.
+However, one characteristics that stand out throughout the entirety of our analysis is the prominence of shooting as the major type of homicide.
+This trend has prevailed from 2004 to 2020.
+We could conclude that shooting is the predominant cause of homicide over time.
 
-We do not observe a general increase or decrease of homicide cases throughout time.
-However, afte the number of homicide cases peaked in 2018, the cases are experiencing a decline.
+We do not observe an apparent increase or decrease of homicide cases throughout time.
+However, after the number of homicide cases peaked in 2018, the cases are experiencing a decline.
 Cases caused by shooting are also in decline during the same period.
@@
 
-Our analysis would be more comprehensive if there are more detailed data about more relevant characteristics about homicides and more cases of homicide.
-But due to the scope of our dataset and the nature of homicide occurrences being generally rare compared to other research topics, it is very hard to give out a firm statement such as a certain factor causes more homicide incidences or homicide incidences are more likely to happen on a certain dates.
+Our analysis would be more comprehensive if there are more detailed data on relevant characteristics and a larger sample of homicide cases.
+But due to the scope of our dataset and the rarity of homicide occurrences, it is very difficult to offer a firm statement regarding a specific factor causing more homicide incidences or predicting homicide incidences are more likely to happen on a certain dates.
 
@@
 
-## References
+# References

### 2024-01-25 outputs/paper/paper.qmd
Made changes based on Rohan's comments and peer review.

@@
 Using data from OpenDataToronto, this paper aims to investigate all recorded homicide cases in Toronto from 2004-2020.
-By exploring data encompassing all Toronto neighbourhoods and focusing those with the highest incidence of homicide cases, I found that shooting is the predominant form of homicide.
+By exploring data encompassing all Toronto neighbourhoods and focusing on those with the highest incidence of homicide cases, I found that shooting is the predominant form of homicide.
 This paper reveals that homicide occurrences in Toronto, in general, do not follow a noticeable temporal pattern, as there is no correlation between the frequency of homicides and temporal variables."
@@
 This research seeks to analyze the spatial and temporal distribution of Toronto’s homicide cases.
-By investigating whether specific times or certain areas are prone to be associated with homicide incidents, we want to shed light on the dynamics of and think about addressing and preventing future prevention for homicide cases.
+By investigating whether specific times or certain areas are prone to be associated with homicide incidents, we want to shed light on the dynamics of homicide occurrences and think about addressing and preventing future prevention for homicide cases.
+My findings showed that homicide occurrences do not adhere to any specific pattern, and therefore are inherently unpredictable.
+However, the likelihood of homicide resulting from shooting is significantly higher compared to other methods.
+This underscores the need for policymakers and law enforcement agencies to take more effective and preventative measures into aspects in firearms.
 
@@
 Data was cleaned and analyzed using the open source statistical programming language R [@citer], with additional support from 'tidyverse' [@rtidyverse], 'dplyr' [@rdplyr], 'janitor'[@rjanitor], 'readr' [@rreadr], 'stringr' [@rstringr], 'lubridate' [@rlubridate], 'ggplot2' [@rggplot2], 'knitr' [@rknitr], 'gt'[@rgt].
-Further details about data extraction and analyzation processes will be discussed in the following sections below.
+Further details about data extraction and analysis processes will be discussed in the following sections below.
 
-This dataset includes a thorough list of alongside every homicide case in Toronto from the year 2004-2020.
+This dataset includes a thorough list of every homicide case in Toronto from the year 2004 to 2020 summing up to 1167 cases.
 I cleaned the dataset and selected variables that are relevant to my analysis of homicide cases.
 My variables of interest include the year the offence occurred, type of homicide (shooting, stabbing, other), date of offence, and identifier of neighbourhood using Toronto’s 140 neighbourhood structure.
-In addition to these variables, I also constructed three more variables: day of the week when the offence occurred (Monday-Sunday), season when the offence occurred (fall, winter, spring, summer), and count of homicide cases by year to further aid our data analysis [@tbl-cleaned_data].
+In addition to these variables, I also constructed three more variables: the day of the week when the offence occurred (Monday-Sunday), the season when the offence occurred (fall, winter, spring, summer), and the count of homicide cases by year to further aid our data analysis [@tbl-cleaned_data].
 This dataset was last refreshed on January 11, 2024.
@@
 I listed out the number of homicide cases each year in @tbl-homicide_count and created the summary statistics table detailing the homicide count per year.
-I found the year 2011 has the lowest cases of homicide, totaling only 51, while 2018 with the total of 97 has the most number of homicide cases.
+I found the year 2011 has the lowest number of cases of homicide, totalling only 51, while 2018 with a total of 97 has the most number of homicide cases.
 
@@
 ```
-Of all the homicide cases analyzed, the cases related to shooting is significantly higher than other type of homicides.
-Over the years of under examination, shooting incidents accounted for more than 600 cases of homicides, surpassing stabbing, which accounted for over 250 cases, and all other types combined which takes up around 300 cases (@fig-distribution_by_type).
+Of all the homicide cases analyzed, the cases related to shooting are significantly higher than other types of homicides.
+Over the years under examination, shooting incidents accounted for more than 608 cases of homicides, surpassing stabbing, which accounted for 260 cases, and all other types combined which took up around 298 cases (@tbl-distribution_by_type).
 
-```{r,out.width='70%', echo=FALSE}
+```{r}
 #| message: false
 #| echo: false
-#| label: fig-distribution_by_type
-#| fig-cap: Distribution of Homicide Types
+#| label: tbl-distribution_by_type
+#| tbl-cap: Distribution of Homicide Types
 
-ggplot(cleaned_homicide_data, aes(x = homicide_type)) +
-  geom_bar(fill = "skyblue", color = "black") +
-  labs(title = "Distribution of Homicide Types",
-       x = "Homicide Types",
-       y = "Count") +
-  theme_minimal()
+# Assuming 'homicide_type' is the column containing homicide types
+homicide_types_table <- cleaned_homicide_data %>%
+  group_by(homicide_type) %>%
+  summarise(total_cases = n())
 
+# Print the table using kable
+kable(homicide_types_table, format = "html", caption = "Number of Homicides by Homicide Types") 
 ```
@@
 With no obvious trend evident, I do not observe a consistent increase or decrease in cases over time.
-However, it is noteworthy that the number homicide cases are decreasing after it peaked in 2018.
+However, it is noteworthy that the number of homicide cases has decreased after it peaked in 2018.
 
@@
 Shooting cases always prevail over other types of homicide cases even during their lowest year.
-We conclude that there is no evident of clear increase or decrease in homicides across time.
+We conclude that there is no evidence of a clear increase or decrease in homicides across time.
 
@@
 
-We want to identify whether if there are specific times in the year when homicides are more likely to occur.
+We want to identify whether there are specific times in the year when homicides are more likely to occur.
 To explore this topic, we generated @fig-case_by_week that displays the homicide counts by each day of the week.
@@
 On the other hand, Thursday has the lowest count, with only 124 cases.
-This analyzation suggests that homicides are slightly more likely to occur on weekends.
+This analysis suggests that homicides are slightly more likely to occur on weekends.
 
@@
                               "Thursday" = "Thu.", "Friday" = "Fri.", "Saturday" = "Sat.", "Sunday" = "Sun.")) +
+  labs(fill = "Day of the Week") +
   theme_minimal()
@@
 This bar plot illustrates the distribution of homicide cases in Toronto's 140 neighbourhoods.
-This visualization enables the identification of patterns or areas of concerns in the city.
+This visualization enables the identification of patterns or areas of concern in the city.
+The spectrum of homicide cases in different neighbourhoods spans from 0 to as many as 31 cases.
 While most neighbourhoods report fewer than 10 cases, it is noteworthy that several neighbourhoods stand out with significantly higher homicide incidences compared to the rest.
@@
 Having observed the significantly higher cases of homicide in specific neighbourhoods, I selected the five neighbourhoods with the highest number of homicide cases-neighbourhoods 73, 25, 132, 26, and 2 with neighbourhood 73 having the highest number of cases of 31 and neighbourhood 2 with 28 cases.
-Through targeted analysis, we can discern if there is specific patterns or trends in the occurrence of homicide incidences in these neighbourhoods and generalize the conclusion in broader contexts.
+Through targeted analysis, we can discern if there are specific patterns or trends in the occurrence of homicide incidences in these neighbourhoods and generalize the conclusion in broader contexts.
 
-In order to observe the relationship between various relating factors and homicide cases, I examined the correlation of the occurrence of each case with each day of the week, season, and type of homicide (@tbl-top5_dayofweek, @tbl-top5_season, @tbl-top5_homicidetype).
-Interestingly, we do not observe a specific trend in the relation between homicide cases and day of the week.
-Similar to our earlier analysis on @fig-case_by_season, the correlation between the number of homicide cases and with a specific day of the week or the weekends are also not noticeable.
-We could attribute this result to limited sample size, given the relatively small number of homicide cases in the city compared to other crimes.
-However, we do notice that just like our earlier results, the amount of homicide caused by shooting is consistently higher than any other methods.
-This is reaffirms that shooting is the predominant type of homicide in Toronto.
+In order to observe the relationship between various related factors and homicide cases, I examined the correlation of the occurrence of each case with each day of the week, season, and type of homicide (@tbl-top5_dayofweek, @tbl-top5_season, @tbl-top5_homicidetype).
+Interestingly, we do not observe a specific trend in the relation between homicide cases and the day of the week.
+Similar to our earlier analysis on @fig-case_by_season, the correlation between the number of homicide cases and a specific day of the week or the weekends is also not noticeable.
+We could attribute this result to a limited sample size, given the relatively small number of homicide cases in the city compared to other crimes.
+However, we do notice that just like our earlier results, the amount of homicide caused by shooting is consistently higher than any other method.
+This reaffirms that shooting is the predominant type of homicide in Toronto.
 
@@
 
-The outcome of our overall analysis of homicide cases by season and by each day of the week throughout Toronto generally align with our specific focus into the top five neighbourhoods with the highest homicide cases.
-We discern no significant correlation between homicidece occurrences and particular timings.
-Homicides may have a slightly higher likelihood in occurring on weekends compared to weekdays but there is no distinct pattern of homicide cases in a certain season.
+The outcome of our overall analysis of homicide cases by season and by each day of the week throughout Toronto generally aligns with our specific focus on the top five neighbourhoods with the highest homicide cases.
+We discern no significant correlation between homicide occurrences and particular timings.
+Homicides may have a slightly higher likelihood of occurring on weekends compared to weekdays but there is no distinct pattern of homicide cases in a certain season.
 
-However, one characteristics that stand out throughout the entirety of our analysis is the prominence of shooting as the major type of homicide.
+However, one characteristic that stands out throughout the entirety of our analysis is the prominence of shooting as the major type of homicide.
 This trend has prevailed from 2004 to 2020.
 We could conclude that shooting is the predominant cause of homicide over time.
+This finding emphasizes the need for potential targeted interventions specifically addressing challenges associated with gun violence in Toronto.
 
-We do not observe an apparent increase or decrease of homicide cases throughout time.
+We do not observe an apparent increase or decrease in homicide cases throughout time.
 However, after the number of homicide cases peaked in 2018, the cases are experiencing a decline.
-Cases caused by shooting are also in decline during the same period.
+Cases caused by shootings are also in decline during the same period.
 This decline may be associated with how Toronto gun violence has dropped by 30% over the past five years (@dropingun).
 
-Our analysis would be more comprehensive if there are more detailed data on relevant characteristics and a larger sample of homicide cases.
-But due to the scope of our dataset and the rarity of homicide occurrences, it is very difficult to offer a firm statement regarding a specific factor causing more homicide incidences or predicting homicide incidences are more likely to happen on a certain dates.
+Our analysis would be more comprehensive if there were more detailed data on relevant characteristics and a larger sample of homicide cases.
+However, due to the scope of our dataset and the rarity of homicide occurrences, it is very difficult to offer a firm statement regarding a specific factor causing more homicide incidences or predicting homicide incidences are more likely to happen on certain dates.
+Despite these limitations, our analysis provides a valuable foundation for future research targeted in addressing homicide dynamics in Toronto.
 

### 2024-01-25 scripts/00-simulate_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
-# Purpose: Simulates homicide data
-# Author: Hannah Yu [...UPDATE THIS...]
-# Date: 17 January 2024 [...UPDATE THIS...]
+# Purpose: Simulates Homicide Data
+# Author: Hannah Yu
+# Date: 17 January 2024 
 # Contact: realhannah.yu@mail.utoronto.ca 
 # License: MIT
 # Pre-requisites: None
@@ -23,7 +23,7 @@ library(ggplot2)
 # The year, month, and day variables contain values within their specified ranges.
 # The weekday variable only include the days of the week from Monday to Sunday.
 # The homicide_type variable only include the following categories: "Shooting," "Stabbing," and "Other." 
-# The neighborhood_number variable include integers ranging from 1 to 158.
+# The neighborhood_number variable include integers ranging from 1 to 140.
 # Columns: year, month, day, weekday, homicide_type, neighborhood_number
 
 #### Simulate Toronto Homicide Data ####
@@ -36,8 +36,8 @@ num_obs <- 1398
 simulated_data <- 
   tibble(
     year = sample(2004:2020, num_obs, replace = TRUE),
-    month = sample(1:12, num_obs, replace = TRUE),
-    day = sample(1:31, num_obs, replace = TRUE),
+    #month = sample(1:12, num_obs, replace = TRUE),
+    #day = sample(1:31, num_obs, replace = TRUE),
     weekday = sample(c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday",
                        "Friday", "Saturday"), num_obs, replace = TRUE),
     homicide_type = sample(c("Shooting", "Stabbing", "Other"),  num_obs,
@@ -45,6 +45,22 @@ simulated_data <-
     neighborhood_number = sample(1:140, num_obs, replace = TRUE),
 )
 
+simulated_data <- 
+  tibble(
+    year = sample(2004:2020, num_obs, replace = TRUE),
+    weekday = sample(c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday",
+                       "Friday", "Saturday"), num_obs, replace = TRUE),
+    homicide_type = sample(c("Shooting", "Stabbing", "Other"),  num_obs,
+                           replace = TRUE),
+    neighborhood_number = sample(1:140, num_obs, replace = TRUE),
+  )
+
+# Create a date column using lubridate
+simulated_data <-
+  simulated_data |>
[…]

### 2024-01-25 scripts/02-data_cleaning.R

@@ -39,7 +39,6 @@ cleaned_homicide_data <-
   cleaned_homicide_data |>
   select(
     occurrence_year, 
-#    division,
     homicide_type,
     occurrence_date,
     hood_id
@@ -50,8 +49,6 @@ cleaned_homicide_data <-
 cleaned_homicide_data <-
   cleaned_homicide_data |>
   mutate(day_of_week = weekdays(occurrence_date))
-         # weekend = ifelse(day_of_week %in% c("Saturday", "Sunday"), TRUE,
-         #                  FALSE))
                         
 # find the season when the homicide occurred #
 cleaned_homicide_data <-

### 2024-01-25 outputs/paper/paper.qmd
Some final edits

@@
 However, the likelihood of homicide resulting from shooting is significantly higher compared to other methods.
-This underscores the need for policymakers and law enforcement agencies to take more effective and preventative measures into aspects in firearms.
+This underscores the need for policymakers and law enforcement agencies to take more effective and preventative measures into aspects of firearms.
 
@@
        y = "Number of Homicides") +
+  labs(fill = "Season") +
   theme_minimal()

### 2024-01-25 scripts/00-simulate_data.R

@@ -24,26 +24,16 @@ library(ggplot2)
 # The weekday variable only include the days of the week from Monday to Sunday.
 # The homicide_type variable only include the following categories: "Shooting," "Stabbing," and "Other." 
 # The neighborhood_number variable include integers ranging from 1 to 140.
-# Columns: year, month, day, weekday, homicide_type, neighborhood_number
+# Columns: year, weekday, homicide_type, neighborhood_number, month, day, date
 
 #### Simulate Toronto Homicide Data ####
 #based on code from: https://tellingstorieswithdata.com/02-drinking_from_a_fire_hose.html#simulate
 
 set.seed(853)
 
-num_obs <- 1398
+num_obs <- 1167
 
-simulated_data <- 
-  tibble(
-    year = sample(2004:2020, num_obs, replace = TRUE),
-    #month = sample(1:12, num_obs, replace = TRUE),
-    #day = sample(1:31, num_obs, replace = TRUE),
-    weekday = sample(c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday",
-                       "Friday", "Saturday"), num_obs, replace = TRUE),
-    homicide_type = sample(c("Shooting", "Stabbing", "Other"),  num_obs,
-                           replace = TRUE),
-    neighborhood_number = sample(1:140, num_obs, replace = TRUE),
-)
+# Create year, weekday, homicide_type, and neighborhood_number columns
 
 simulated_data <- 
   tibble(
@@ -67,8 +57,7 @@ simulated_data <-
 print(simulated_data)
 
 
-#Bar Plot for Homicide Type
-
+# Bar Plot for Homicide Type
 
 ggplot(simulated_data, aes(x = homicide_type, fill = homicide_type)) +
   geom_bar() +
@@ -76,7 +65,7 @@ ggplot(simulated_data, aes(x = homicide_type, fill = homicide_type)) +
        y = "Count") +
   theme_minimal()
 
-#Histogram for Neighborhood Number
+# Histogram for Neighborhood Number
 
 ggplot(simulated_data, aes(x = neighborhood_number)) +
   geom_histogram(binwidth = 1, fill = "blue", color = "black") +

### 2024-01-25 scripts/01-download_data.R

@@ -11,10 +11,8 @@
 library(opendatatoronto)
 library(tidyverse)
 library(dplyr)
-# [...UPDATE THIS...]
 
 #### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
 
 raw_homicide_data <-
 
@@ -25,9 +23,6 @@ raw_homicide_data <-
 
 
 #### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-#write_csv(the_raw_data, "inputs/data/raw_data.csv") 
 
 write_csv(
   x = raw_homicide_data,

### 2024-01-25 scripts/02-data_cleaning.R

@@ -26,15 +26,15 @@ raw_homicide_data <-
     show_col_type = FALSE
   )
 
-# Make the names easier to type 
+# Make the names easier to type #
 cleaned_homicide_data <-
   clean_names(raw_homicide_data)
 
-# Have a look at the first six rows
+# Have a look at the first six rows #
 head(cleaned_homicide_data)
 
 
-# select columns of interest #
+# Select columns of interest 
 cleaned_homicide_data <-
   cleaned_homicide_data |>
   select(
@@ -45,7 +45,7 @@ cleaned_homicide_data <-
   )
 
 
-# find the day of the week of the homicide and determine if it is a weekend #
+# Find the day of the week of the homicide and determine if it is a weekend #
 cleaned_homicide_data <-
   cleaned_homicide_data |>
   mutate(day_of_week = weekdays(occurrence_date))
